Programmatic Generation of R Markdown Documents with \pkg{listdown}

### 2020-04-28 inst/r-journal/listdown-article.tex
Major bigtime revision.

Since computational components are, by definition, computationally
derived objects and R Markdown is a well-defined standard it is possible
to programmatically create R Markdown documents with computational
components. This is the focus of this paper. However, before proceeding
down this avenue, we would like to highlight two fundamental limitations
to automated R Markdown document construction. First and foremost,
without narrative components a document has no context. Quantitative
analyses require research questions, hypotheses, reviews,
interpretations, and conclusions. Computational components are necessary
but not sufficient for constructing an analysis. Second, it is not
possible to construct computational components for an arbitary set of
distinct analyses. An analysis itself has a context and it is built with
a set of assumptions and goals. Computational components are constructed
for a narrow class of problems.

However, this is not to say the programmatic creation of R Markdown
documents is unwarranted or not useful. In fact it has at least two
appealing characteristics. The first is convenience. \textless FILL THIS
IN - TALK ABOUT REUSE TO MOTIVATE LARGER PIPELINES\textgreater. The
second is that for fixed analysis pipelines processing
uniformly-formatted but domain-distinct data, automated document
generation enforces uniformity across those domains being serviced. .

The \pkg{listdown} package provides functions to programmatically create
R Markdown files from named lists. It is intended for data analysis
pipelines where the presentation of the results is separated from their
creation. For this use case, a data processing (or analysis) is
performed and the results are provided in a single named list, organized
heirarchically. With the list and a \pkg{listdown} object a workflowr
\citep{blischak2019}, pdf, word, or html page. List element names denote
sections, subsections, subsubsection, etc. and the list elements contain
the data structure to be presented including graphs and tables. The goal
of the package is not to provide a finished, readable document. It is to
provide a document with all tables and visualization that will appear
(\emph{computational} components). This serves as a starting point from
which a user can organize outputs, describe a study, discuss results,
and provide conclusions (\emph{narrative} components).

@@

The package itself is relatively simple with 6 distinct methods that can
be easily incorporated into existing analysis pipelines for
automatically creating documents that can be used for data exploration
and reviewing analysis results as well as a starting point for a more
formal write up. These methods include:

\begin{itemize}
\item{\bf ld\_chunk\_opts() }{- apply chunk options to a presentation object}
\item{\bf ld\_ioslides\_header() }{- create an ioslides presentation header}
\item{\bf ld\_make\_chunks() }{- write a listdown object to a string}
\item{\bf ld\_workflowr\_header() }{- create a worflowr header}
\item{\bf ld\_rmarkdown\_header() }{- create an R Markdown header}
\item{\bf listdown() }{- create a listdown object to create an R Markdown document}
\item{\bf to\_data\_tree() }{- create a data.tree object from a list of computational components}
\end{itemize}

@@

Creating a document from the \texttt{computational\_components} will
require two steps. First, we will create a \texttt{listdown} object that
specifies how the \texttt{computational\_components} object will be
loaded into the document, which libraries and code needs to be included,
the options for the R chunks, and how the list elements will be
presented in the output R markdown document.

@@

ld <- listdown(load_cc_expr = readRDS("comp-comp.rds"), # The expression to load pres_list in the .Rmd.
               package = "ggplot2")                     # The packges needed to render plots.
\end{Sinput}
\end{Schunk}

The \texttt{ld} object, along with the computational components in the
\texttt{comp-comp.rds} file are sufficient to to create the sections,
subsections, and R chunks of a document. The only other thing requires
to create the document is the header. The listdown package currently
supports regular R Markdown, workflowr, and ioslides headers. A complete
document can then be written to the console using the code shown below.
It could easily be written to a file for rendering using the
\texttt{writeLines()} function, for example.

\begin{Schunk}
\begin{Sinput}
doc <- c(
  ld_rmarkdown_header("Anscombe's Quartet",
                     author = "Francis Anscombe",
                     date = "1973"),
  ld_make_chunks(ld))

cat(paste(doc, collapse = "\n"))
\end{Sinput}
\begin{Soutput}
#> ---
#> title: "Anscombe's Quartet"
#> author: "Francis Anscombe"
#> date: "1973"
#> output: html_document
#> ---
#> 
#> ```{r}
#> library(ggplot2)
#> 
#> cc_list <- readRDS("comp-comp.rds")
#> ```
#> 
#> # Linear
#> 
#> ```{r}
#> identity(cc_list[[1]])
#> ```
#> 
#> # Non Linear
#> 
#> ```{r}
#> identity(cc_list[[2]])
#> ```
#> 
#> # Outlier Vertical
#> 
#> ```{r}
#> identity(cc_list[[3]])
#> ```
#> 
#> # Outlier Horizontal
#> 
#> ```{r}
#> identity(cc_list[[4]])
#> ```
\end{Soutput}
\end{Schunk}

@@

The first example is simple in part because the ggplot objects both
contain the data we want to display and, at the same time, provide the
mechanism for presenting them - rendering them in a graph. However, this
is not always the case. The objects being stored in the list of
computational components may not translate directly to the presentation
in a document. In these cases a function is needed that takes the list
component and returns an object to be displayed. For example, suppose
that, along with showing graphs from the Anscombe Quartet, we would like
to include the data themselves. We could add the data to the
\texttt{computational\_components} list and then create the document
with:

\begin{Schunk}
\begin{Sinput}
computational_components$Data <- anscombe
saveRDS(computational_components, "comp-comp.rds")
cat(paste(ld_make_chunks(ld)[32:36], collapse = "\n"))
\end{Sinput}
\begin{Soutput}
#> # Data
#> 
#> ```{r echo=FALSE}
#> identity(cc_list[[5]])
#> ```
\end{Soutput}
\end{Schunk}

@@

In \pkg{listdown} a function or method that implments the presentation
of a computational component is referred to as a \emph{decorator} since
if follows the classic decorator pattern described in \citet{gamma1995}.
A decorator takes the element that will be presented as an argument and
returns an object for presentation in the output directory. A decorator
is specified using the \texttt{decorator} parameter of the
\texttt{listdown()} function using a named list where the name
corresponds to the type and the element correspond to the function or
method that will decorate an object of that type. For example, the
\texttt{anscombe} data set can be decorated with the
\texttt{DT::datatable()} function \citep{xie2020} as:

\begin{Schunk}
\begin{Sinput}
ld <- listdown(load_cc_expr = readRDS("comp-comp.rds"), 
               package = c("ggplot2", "DT"),
               decorator = list(data.frame = datatable))
cat(paste(ld_make_chunks(ld)[33:37], collapse = "\n"))
\end{Sinput}
\begin{Soutput}
#> # Data
#> 
#> ```{r}
#> datatable(cc_list[[5]])
#> ```
\end{Soutput}
\end{Schunk}

List names in the \texttt{decorator} argument provide a key to which a
function or method is mapped. The underlying decorator resolution is
implemented for a given computational component by going through
decorator names sequentially to see if the component inherits from the
name using the \texttt{inherits()} function. The function or method is
selected from the corresponding name which the element first inherits
from. This means that when customizing the presentation of objects that
inherit from a common class, the more abstract classes should appear at
the end of the list. This will ensure that specialized classes will be
encountered first in the resolution process.

@@

# Create the dendrogram.
ld_cc_dendro(comp_comp2)
\end{Sinput}
\begin{Soutput}
#> 
#>  comp_comp2
#>   |-- Iris
#>   |  °-- object of type(s):data.frame
#>   °-- Sepal.Length
#>    |-- Sepal.Width
#>    |  °-- object of type(s):gg ggplot
#>    |-- Petal.Length
#>    |  °-- object of type(s):gg ggplot
#>    °-- Colored
#>     |-- Sepal.Width
#>     |  °-- object of type(s):gg ggplot
#>     °-- Petal.Length
#>        °-- object of type(s):gg ggplot 
#> 
\end{Soutput}
\end{Schunk}

@@

\hypertarget{advanced-usage}{%
\section{Advanced usage}\label{advanced-usage}}

Along with the basic usage, the \pkg{listdown} package supports methods
to further customize the presentation. Two specific areas of
customization are addressed. The first is where custom code is required
either to modify decorators or even define the decorators. In this case,
we would to add a small portion of code, immediately after loading
required libraries. In the second case, rather than changing the R code
inside of a chunk, we would like to change the options for an individual
chunk. Both of these features are supported in the package and are
presented below.

@@

\begin{Schunk}
\begin{Sinput}
saveRDS(comp_comp2, "comp-comp2.rds")
ld <- listdown(load_cc_expr = readRDS("comp-comp2.rds"),
               package = c("ggplot2", "DT", "purrr"),
               decorator = list(ggplot = identity,
                                data.frame = datatable),
               init_expr = {
                 datatable <- partial(datatable,
                                      options = list(dom = 't'))
                 },
               echo = FALSE)

cat(paste(ld_make_chunks(ld)[1:13], collapse = "\n"))
\end{Sinput}
\begin{Soutput}
#> 
#> ```{r echo=FALSE}
#> library(ggplot2)
#> library(DT)
#> library(purrr)
#> 
#> cc_list <- readRDS("comp-comp2.rds")
#> ```
#> 
#> ```{r echo=FALSE}
#> datatable <- partial(datatable, options = list(dom = "t"))
#> ```
\end{Soutput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
comp_comp2$Iris <- ld_chunk_opts(comp_comp2$Iris, echo = TRUE)
saveRDS(comp_comp2, "comp-comp2.rds")
cat(paste(ld_make_chunks(ld)[1:19], collapse = "\n"))
\end{Sinput}
\begin{Soutput}
#> 
#> ```{r echo=FALSE}
#> library(ggplot2)
#> library(DT)
#> library(purrr)
#> 
#> cc_list <- readRDS("comp-comp2.rds")
#> ```
#> 
#> ```{r echo=FALSE}
#> datatable <- partial(datatable, options = list(dom = "t"))
#> ```
#> 
#> # Iris
#> 
#> ```{r echo = TRUE}
#> datatable(cc_list[[1]])
#> ```
\end{Soutput}
\end{Schunk}

\hypertarget{conclusion}{%
\section{Conclusion}\label{conclusion}}

>>>>

Since computational components are, by definition, computationally
derived objects and R Markdown is a well-defined standard it is possible
to programmatically create R Markdown documents with computational
components. This is the focus of this paper. However, before proceeding
down this avenue, we would like to highlight two fundamental limitations
to automated R Markdown document construction. First and foremost,
without narrative components a document has no context. Quantitative
analyses require research questions, hypotheses, reviews,
interpretations, and conclusions. Computational components are necessary
but not sufficient for constructing an analysis. Second, it is not
possible to construct computational components for an arbitary analyses.
Analyses themselves have context and are built with a set of assumptions
and goals. Computational components are constructed for a narrow class
of problems.

However, this is not to say the programmatic creation of R Markdown
documents is unwarranted or not useful. In fact, it has at least two
appealing characteristics. The first is that it allows us to distinguish
the presentation of analytical results from other steps in a data
science or data processing pipeline. These other steps including
cleaning and analysis often require their own environment and
configuration with requirements very different than the computational
needs of creating a presentation. By separating these components each
can be developed independently. At the same time, by specifying that the
output of those objects are organized hierarchically, we are essentially
creating a contract by which processed data can be passed to the
presentation as well as providing a way to organize the results. The
second reason for programmatic creation of R Markdown documents is
convenience. In collaborative environments, especially in the early
stages, large numbers of graphs and tables are generated and discussed.
By collecting these artifacts hierarchicaly as a list, they can be
quickly iterated upon and restructured to more clearly present the data
without needing to spend time on the creation of the presentation
document.

The \pkg{listdown} package provides functions to programmatically create
R Markdown files from named lists. It is intended for data analysis
pipelines where the presentation of the results is separated from their
creation. For this use case, a data processing (or analysis) is
performed and the results are provided in a single named list, organized
heirarchically. With the list and a \pkg{listdown} object a
\pkg{workflowr} \citep{blischak2019}, pdf, word, or html page. List
element names denote sections, subsections, subsubsection, etc. and the
list elements contain the data structure to be presented including
graphs and tables. The goal of the package is not to provide a finished,
readable document. It is to provide a document with all tables and
visualization that will appear (\emph{computational} components). This
serves as a starting point from which a user can organize outputs,
describe a study, discuss results, and provide conclusions
(\emph{narrative} components).

@@

One area where we have found \pkg{listdown} is particuarly useful is in
the reporting and research of clinical trial data. These collaborations
tend to be between (bio)statisticians and clinicians either analyzing
past trial data to formulate a new trial or in trial monitoring where
trial telemetry (enrollment, responses, etc.) is reported and initial
analyses are conveyed to a clinician. The associated presentations
require very little context - clinicians often have a better
understanding of the data collected than the statistician - often
eliminating the need for narrative components. At the same time, a large
number of hierarchical, heterogenous artifacts (tables and multiple
types of plots) need to be conveyed thereby making the manual creation
of R Markdown documents inconvenient.

Furthermore, data processing can be fixed across trials. This is
especially true in the initial stages, which focus on patient
demographics and enrollment. This has made it convenient for our group
to quickly generate standardize and complete reports for multiple trials
concurrently. To date, we have used listdown to report on five clinical
trials, with another two currently in process. Results are disseminated
using the \pkg{workflowr} package, usually with nine tabs conveying
aspects of the data from collection through several different analyses,
and each tab containing approxiately five to thirty tables, plots, or
other artifacts including \pkg{trelliscopejs} \citep{trelliscopejs}
environments which may hold hundreds of graphs. By generating many
presentation artifacts we are able to address data-drive questions and
issues during collaborative sessions and by carefully structuring these
elements we allowing all members to participate in the process.

The \pkg{listdown} package itself is relatively simple with 6 distinct
methods that can be easily incorporated into existing analysis pipelines
for automatically creating documents that can be used for data
exploration and reviewing analysis results as well as a starting point
for a more formal write up. These methods include:

\begin{itemize}
\item{\bf as\_ld\_yml() }{- turn a computational component list into YAML with class information}
\item{\bf ld\_cc\_dendro() }{- create a dendrogram from a list of computational components}
\item{\bf ld\_chunk\_opts() }{- apply chunk options to a presentation object}
\item{\bf ld\_ioslides\_header() }{- create an ioslides presentation header}
\item{\bf ld\_make\_chunks() }{- write a listdown object to a string}
\item{\bf ld\_rmarkdown\_header() }{- create an R Markdown header}
\item{\bf ld\_workflowr\_header() }{- create a worflowr header}
\item{\bf listdown() }{- create a listdown object to create an R Markdown document}
\end{itemize}

@@

Creating a document from the \texttt{computational\_components} will
require two steps. First, we will create a \texttt{listdown} object that
specifies how the \texttt{computational\_components} object will be
loaded into the document, which libraries and code needs to be included,
and how the list elements will be presented in the output R markdown
document.

@@

ld <- listdown(load_cc_expr = readRDS("comp-comp.rds"),
               package = "ggplot2")
\end{Sinput}
\end{Schunk}

The \texttt{ld} object, along with the computational components in the
\texttt{comp-comp.rds} file are sufficient to to create the sections,
subsections, and R chunks of a document. The only other thing requires
to create the document is the header. The listdown package currently
supports regular R Markdown and \pkg{workflowr} as \code{yml} objects
from the \pkg{yaml} package \citep{yaml}. These objects are stored as
named lists in R and are easily modified to accomodate document
parameters. A complete document can then be written to the console using
the code shown below. It could easily be written to a file for rendering
using the \texttt{writeLines()} as in the following example.

\begin{Schunk}
\begin{Sinput}
doc <- c(
  as.character(
    ld_rmarkdown_header("Anscombe's Quartet",
                        author = "Francis Anscombe",
                        date = "1973")),
  ld_make_chunks(ld))

cat(paste(doc, collapse = "\n"))
\end{Sinput}
\begin{Soutput}
#> ---
#> title: Anscombe's Quartet
#> author: Francis Anscombe
#> date: '1973'
#> output: html_document
#> ---
#> 
#> ```{r}
#> library(ggplot2)
#> 
#> cc_list <- readRDS("comp-comp.rds")
#> ```
#> 
#> # Linear
#> 
#> ```{r}
#> cc_list$Linear
#> ```
#> 
#> # Non Linear
#> 
#> ```{r}
#> cc_list$`Non Linear`
#> ```
#> 
#> # Outlier Vertical
#> 
#> ```{r}
#> cc_list$`Outlier Vertical`
#> ```
#> 
#> # Outlier Horizontal
#> 
#> ```{r}
#> cc_list$`Outlier Horizontal`
#> ```
\end{Soutput}
\end{Schunk}

@@

The first example is simple in part because the ggplot objects both
contain the data we want to display and, at the same time, provide the
mechanism for presenting them - rendering them in a graph. However, this
is not always the case. The objects being stored in the list of
computational components may not translate directly to the presentation
in a document. In these cases, a function is needed that takes the list
component and returns an object to be displayed. For example, suppose
that, along with showing graphs from the Anscombe Quartet, we would like
to include the data themselves. We could add the data to the
\texttt{computational\_components} list and then create the document
with:

\begin{Schunk}
\begin{Sinput}
computational_components$Data <- anscombe
saveRDS(computational_components, "comp-comp.rds")
cat(paste(ld_make_chunks(ld)[32:36], collapse = "\n"))
\end{Sinput}
\begin{Soutput}
#> # Data
#> 
#> ```{r echo=FALSE}
#> cc_list$Data
#> ```
\end{Soutput}
\end{Schunk}

@@

In \pkg{listdown}, a function or method that implements the presentation
of a computational component is referred to as a \emph{decorator} since
if follows the classic decorator pattern described in \citet{gamma1995}.
A decorator takes the element that will be presented as an argument and
returns an object for presentation in the output directory. A decorator
is specified using the \texttt{decorator} parameter of the
\texttt{listdown()} function using a named list where the name
corresponds to the type and the element correspond to the function or
method that will decorate an object of that type. For example, the
\texttt{anscombe} data set can be decorated with the
\texttt{DT::datatable()} function \citep{xie2020} as:

\begin{Schunk}
\begin{Sinput}
ld <- listdown(load_cc_expr = readRDS("comp-comp.rds"), 
               package = c("ggplot2", "DT"),
               decorator = list(data.frame = datatable))
cat(paste(ld_make_chunks(ld)[33:37], collapse = "\n"))
\end{Sinput}
\begin{Soutput}
#> # Data
#> 
#> ```{r}
#> datatable(cc_list$Data)
#> ```
\end{Soutput}
\end{Schunk}

List names in the \texttt{decorator} argument provide a key to which a
function or method is mapped. The underlying decorator resolution is
implemented for a given computational component by going through
decorator names sequentially to see if the component inherits from the
name using the \texttt{inherits()} function. The function or method is
selected from the corresponding name which the element first inherits
from. This means that when customizing the presentation of objects that
inherit from a common class, the more abstract classes should appear at
the end of the list. This will ensure that specialized classes will be
encountered first in the resolution process. It should be noted that an
object's type is first checked against the decorator name list and then
checked to see if it is a list. This allows a user to both decorate a
list and retain \texttt{"list"} in its class attributes.

@@

# Create the dendrogram.
ld_cc_dendro(comp_comp2)
\end{Sinput}
\begin{Soutput}
#> 
#>  comp_comp2
#>   |-- Iris
#>   |  o-- object of type(s):data.frame
#>   o-- Sepal.Length
#>    |-- Sepal.Width
#>    |  o-- object of type(s):gg ggplot
#>    |-- Petal.Length
#>    |  o-- object of type(s):gg ggplot
#>    o-- Colored
#>     |-- Sepal.Width
#>     |  o-- object of type(s):gg ggplot
#>     o-- Petal.Length
#>        o-- object of type(s):gg ggplot 
#> 
\end{Soutput}
\end{Schunk}

@@

\hypertarget{r-code-chunk-customization}{%
\section{R code chunk customization}\label{r-code-chunk-customization}}

Along with the basic usage, the \pkg{listdown} package supports two
other capabilities to further customize the presentation. The first is
where custom code is required either to modify decorators or even define
a decorators. In this case, we would to add a small portion of code,
immediately after loading required libraries. In the second case, rather
than changing the R code inside of a chunk, we would like to change the
options for an individual chunk. Both of these features are supported in
the package and are presented below.

@@

The \texttt{listdown()} function facilitates the insertion of
initialization code through the \texttt{init\_expr} argument. Code is
inserted immediately after the libraries are loaded in the R Markdown
document. In general, it is suggested that the number of initial
expressions be kept small so that the R Markdown document is easy to
read. If a large number of functions are required by the target R
Markdown document then they can be put into a file and sourced using the
initial expression. As an example, suppose we are creating an html
document and presenting data using the \texttt{datatable()} function.
However, we do not want to include the search capabilities provided.
This can be easily accomplished by creating a new function,
\texttt{datatable\_no\_search()}, created using the \texttt{partial()}
function \cite{purrr} to partially apply
\texttt{list(dom\ =\ \textquotesingle{}t\textquotesingle{})} to the
\texttt{options} argument of \texttt{datatable}.

\begin{Schunk}
\begin{Sinput}
saveRDS(comp_comp2, "comp-comp2.rds")
ld <- listdown(load_cc_expr = readRDS("comp-comp2.rds"),
               package = c("ggplot2", "DT", "purrr"),
               decorator = list(ggplot = identity,
                                data.frame = datatable_no_search),
               init_expr = {
                 datatable_no_search <- partial(datatable,
                                                options = list(dom = 't'))
                 },
               echo = FALSE)

cat(paste(ld_make_chunks(ld)[1:13], collapse = "\n"))
\end{Sinput}
\begin{Soutput}
#> 
#> ```{r echo=FALSE}
#> library(ggplot2)
#> library(DT)
#> library(purrr)
#> 
#> cc_list <- readRDS("comp-comp2.rds")
#> ```
#> 
#> ```{r echo=FALSE}
#> 
#> datatable_no_search <- partial(datatable, options = list(dom = "t"))
#> ```
\end{Soutput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
comp_comp2$Iris <- ld_chunk_opts(comp_comp2$Iris, echo = TRUE)
saveRDS(comp_comp2, "comp-comp2.rds")
cat(paste(ld_make_chunks(ld)[1:19], collapse = "\n"))
\end{Sinput}
\begin{Soutput}
#> 
#> ```{r echo=FALSE}
#> library(ggplot2)
#> library(DT)
#> library(purrr)
#> 
#> cc_list <- readRDS("comp-comp2.rds")
#> ```
#> 
#> ```{r echo=FALSE}
#> 
#> datatable_no_search <- partial(datatable, options = list(dom = "t"))
#> ```
#> 
#> # Iris
#> 
#> ```{r echo = TRUE}
#> datatable_no_search(cc_list$Iris)
#> ```
\end{Soutput}
\end{Schunk}

\hypertarget{a-simple-example-reporting-on-the-gtsummarytrial-dataset}{%
\section{\texorpdfstring{A simple example: Reporting on the
\texttt{gtsummary::trial}
dataset}{A simple example: Reporting on the gtsummary::trial dataset}}\label{a-simple-example-reporting-on-the-gtsummarytrial-dataset}}

\label{sect:simple-example}

\begin{Schunk}
\begin{Sinput}
library(gtsummary)
library(dplyr)
library(glue)
library(survival)
library(survminer)
library(rmarkdown)

# Function to create the computational components.
make_surv_cc <- function(trial, treat, surv_cond_chars) {
  # Create Table 1 by treatment
  table_1 <- trial %>%
    tbl_summary(by = all_of(treat)) %>%
    gtsummary::as_flextable()

  # Create survival plots base on survival characteristics
  scs <- lapply(c("1", surv_cond_chars),
                function(sc) {
                  glue("Surv(ttdeath, death) ~ {treat} + {sc}") %>%
                    as.formula() %>%
                    surv_fit(trial) %>%
                    ggsurvplot()
                })
  names(scs) <- c("Overall", tools::toTitleCase(surv_cond_chars))
  list(`Table 1` = table_1, `Survival Plots` = scs)
}

# Create the computational components and write to surv-cc.rds.
make_surv_cc(trial, treat = "trt",
             surv_cond_chars = c("stage", "grade")) %>%
  saveRDS("surv-cc.rds")

# Create the listdown object with appropriate dependencies
# and presentation parameters.
ld_surv <- listdown(load_cc_expr = readRDS("surv-cc.rds"),
                    package = c("gtsummary", "survminer"),
                    decorator = list(ggsurvplot = identity,
                                     tbl_summary = identity),
                    echo = FALSE,
                    message = FALSE,
                    warning = FALSE,
                    fig.width = 7,
                    fig.height = 4.5)

# Write an RMarkdown header and chunks derived from the 
# computational components and listdown object to trial-report.rmd.
writeLines(
  paste(c(
    as.character(ld_rmarkdown_header("Simple Trial Report")),
    ld_make_chunks(ld_surv))),
  "trial-report.rmd")

# Render the RMarkdown file to html.
render("trial-report.rmd", quiet = TRUE, pdf_document(keep_tex = TRUE))  
\end{Sinput}
\end{Schunk}

\hypertarget{using-in-practice-for-clinical-trial-reporting}{%
\section{\texorpdfstring{Using \pkg{listdown} in practice for clinical
trial
reporting}{Using  in practice for clinical trial reporting}}\label{using-in-practice-for-clinical-trial-reporting}}

\pagebreak

\hypertarget{appendix-1-rmarkdown-produced-in-the-simple-example}{%
\section{Appendix 1: RMarkdown produced in the simple
example}\label{appendix-1-rmarkdown-produced-in-the-simple-example}}

\begin{Schunk}
\begin{Sinput}
---
title: Simple Trial Report
output: html_document
---

\```{r echo=FALSE}
library(gtsummary)
library(survminer)

cc_list <- readRDS("surv-cc.rds")
\```

# Table 1

\```{r echo=FALSE}
cc_list$`Table 1`
\```

# Survival Plots

## Overall

\```{r echo=FALSE}
cc_list$`Survival Plots`$Overall
\```

## Stage

\```{r echo=FALSE}
cc_list$`Survival Plots`$Stage
\```

## Grade

\```{r echo=FALSE}
cc_list$`Survival Plots`$Grade
\```
\end{Sinput}
\end{Schunk}

\pagebreak

\hypertarget{appendix-2-the-document-produced-by-the-simple-example}{%
\section{Appendix 2: The document produced by the simple
example}\label{appendix-2-the-document-produced-by-the-simple-example}}
